---
title: "Reproducible Research: Peer Assessment 1"
output:
  html_document:
    keep_md: true
---

```r
library(dplyr)
library(ggplot2)
library(timeDate)

rm(list = ls()) # clear vars
setwd("C:\\dev\\RepData_PeerAssessment1\\")

activity.data <- read.csv(unz("activity.zip", "activity.csv"), colClasses = c("integer", "Date", "integer"))
activity.data$date <- as.POSIXct(activity.data$date)
```

# Calculate the total number of steps taken per day

 ```{r, echo=TRUE}
total.steps <- tapply(activity.data$steps, activity.data$date, FUN = sum, na.rm = TRUE)
daily.steps <- activity.data %>% group_by(date) %>% summarise(Steps = sum(steps))
names(daily.steps) <- c("Date", "steps")
```

# Mean and median number of steps taken each days
 ```{r, echo=TRUE}
mean.daily.steps <- mean(total.steps, na.rm = TRUE)
median.daily.steps <- median(total.steps, na.rm = TRUE)
```

## mean.daily.steps:`r mean.daily.steps`
## median.daily.steps:`r median.daily.steps`

# Histogram of the total number of steps taken each day

```{r steps_per_day, echo=TRUE}
hist(daily.steps$steps, breaks = 11,
     xlab = "number of steps per day",
     main = "Histogram of total steps per day")
abline(v = mean.daily.steps, col = "red", lwd = 1)
abline(v = median.daily.steps, col = "blue", lwd = 1)
legend(x = "topright", legend = c("mean", "median"), col = c("red", "blue"), bty = "n", lwd = 3)
```

## What is the average daily activity pattern?

 ```{r, echo=TRUE}
activity.interval <- activity.data %>% group_by(interval) %>% summarise(mean_steps = mean(steps, na.rm = TRUE))
```

#Time series plot of the average number of steps taken
```{r activity_interval_plot, echo=TRUE}
activity.interval.plot <- ggplot(data = activity.interval, mapping = aes(x = interval, y = mean_steps)) +
    geom_line() + scale_x_continuous("Day Interval", breaks = seq(min(activity.interval$interval),
    max(activity.interval$interval), 100)) + scale_y_continuous("Average Number of Steps") +
ggtitle("Average Number of Steps Taken by Interval")
activity.interval.plot
```

#The 5-minute interval that, on average, contains the maximum number of steps

 ```{r, echo=TRUE}
max.steps = activity.interval[activity.interval$mean_steps==which.max(activity.interval$mean_steps)]
#max.steps:`r max.steps`
```

### Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs)
 ```{r, echo=TRUE}
present.days <- na.omit(activity.data$steps)
missing.days <- length(activity.data$steps) - length(present.days)
# strategy for filling in missing data (just replace with mean)
activity.data$steps[is.na(activity.data$steps)] <- mean.daily.steps
# create a new dataset that is equal to the original dataset but with the missing data filled in.
strategy <- function(x) { ifelse(is.na(x), mean.daily.steps, x) }    
activity.data.no.na <- as.data.frame(lapply(activity.data, FUN = strategy))
```

# repeat first part of assignment using non NA valued data
# Calculate the total number of steps taken per day

 ```{r, echo=TRUE}
total.steps.no.na <- tapply(activity.data.no.na$steps, activity.data.no.na$date, FUN = sum, na.rm = TRUE)
daily.steps.no.na <- total.steps.no.na %>% group_by(date) %>% summarise(Steps = sum(steps))
names(daily.steps.no.na) <- c("Date", "steps")
```

# Calculate and report the mean and median of the total number of steps taken per day

```{r report_mean_median, echo=TRUE}
mean.daily.steps.no.na <- mean(total.steps.no.na, na.rm = TRUE)
median.daily.steps.no.na <- median(total.steps.no.na, na.rm = TRUE)
```

# Histogram of the total number of steps taken each day after missing values are imputed

```{r steps_again, echo=TRUE}
hist(daily.steps.no.na$steps, breaks = 11,
    xlab = "number of steps per day 2",
    main = "Histogram of total steps per day 2")
    abline(v = daily.steps.no.na, col = "red", lwd = 1)
    abline(v = median.daily.steps.no.na, col = "blue", lwd = 1)
    legend(x = "topright", legend = c("mean", "median"), col = c("red", "blue"), bty = "n", lwd = 3)
```

#Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends
#https://cran.r-project.org/web/packages/timeDate/index.html

 ```{r, echo=TRUE}
activity.data$day.type <- "weekday"
activity.data$day.type[isWeekend(activity.data$date)] <- "weekend"
activity.data$day.type <- as.factor(activity.data$day.type)
```

# Make a panel plot containing a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis). 
```{r five_minute_interval, echo=TRUE}
pattern <- aggregate(steps ~ interval + day.type, data = activity.data, mean)
library(lattice)
xyplot(steps ~ interval | day.type, pattern, type = "l", layout = c(1, 2), xlab = "Interval", ylab = "Number of steps")
```

##  Calculate average steps of each 5-minutes interval for weekdays and weekend

```{r weekend_data, echo=TRUE}
weekend <- pattern[which(pattern$day.type == "weekend"),] %>% group_by(day.type) %>% summarise(Steps = sum(steps))
weekday <- activity.data[which(activity.data$day.type == "weekday"),]
```