```{r global_options, include=FALSE}
#suppress the warnings and other messages from showing in the knitted file.
knitr::opts_chunk$set(fig.width = 8, fig.height = 6,eval=TRUE,
                      eval=TRUE, echo=TRUE, warning=FALSE, message=FALSE)
```


# Reproducible Research: Peer Assessment 1

```{r}
library(dplyr)
library(ggplot2)
library(timeDate)
```

# Code for reading in the dataset and/or processing the data
```{r}
rm(list = ls()) # clear vars
setwd("C:\\dev\\RepData_PeerAssessment1\\")
activity.data <- read.csv(unz("activity.zip", "activity.csv"), colClasses = c("integer", "Date", "integer"))
activity.data$date <- as.POSIXct(activity.data$date)
```

# Calculate the total number of steps taken per day
```{r , eval=TRUE, echo=TRUE}
total.steps <- tapply(activity.data$steps, activity.data$date, FUN = sum, na.rm = TRUE)

```

# Mean and median number of steps taken each days
```{r , eval=TRUE, echo=TRUE}
mean.daily.steps <- mean(total.steps, na.rm = TRUE)
median.daily.steps <- median(total.steps, na.rm = TRUE)
```

# mean daily steps = `r mean.daily.steps`
# median daily steps = `r median.daily.steps`

# Histogram of the total number of steps taken each day
```{r five_minute_interval, eval=TRUE, echo=TRUE}
daily.steps <- activity.data %>% group_by(date) %>% summarise(Steps = sum(steps))
names(daily.steps) <- c("Date", "steps")
hist(daily.steps$steps, breaks = 11,
     xlab = "number of steps per day",
     main = "Histogram of total steps per day")
abline(v = mean.daily.steps, col = "red", lwd = 1)
abline(v = median.daily.steps, col = "blue", lwd = 1)
legend(x = "topright", legend = c("mean", "median"), col = c("red", "blue"), bty = "n", lwd = 3)
```

# What is the average daily activity pattern?
```{r , eval=TRUE, echo=TRUE}
activity.interval <- activity.data %>% group_by(interval) %>% summarise(mean_steps = mean(steps, na.rm = TRUE))
activity.interval
```


# Time series plot of the average number of steps taken
```{r , eval=TRUE, echo=TRUE}
activity.interval.plot <- ggplot(data = activity.interval, mapping = aes(x = interval, y = mean_steps)) +
    geom_line() + scale_x_continuous("Day Interval", breaks = seq(min(activity.interval$interval),
    max(activity.interval$interval), 100)) + scale_y_continuous("Average Number of Steps") +
ggtitle("Average Number of Steps Taken by Interval")
activity.interval.plot
```

# The 5-minute interval that, on average, contains the maximum number of steps

```{r , eval=TRUE, echo=TRUE}
max.steps = activity.interval[which.max(activity.interval$mean_steps),]
```

# Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs)

```{r , eval=TRUE, echo=TRUE}
present.days <- na.omit(activity.data$steps)
missing.days <- length(activity.data$steps) - length(present.days)
missing.days
```

# implement a strategy for filling in missing data by replacing NA values with mean)
```{r , eval=TRUE, echo=TRUE}
activity.data$steps[is.na(activity.data$steps)] <- mean.daily.steps
```

# create a new dataset that is equal to the original dataset but with the missing data filled in.
```{r , eval=TRUE, echo=TRUE}
strategy <- function(x) { ifelse(is.na(x), mean.daily.steps, x) }    
activity.data.filled <- as.data.frame(lapply(activity.data, FUN = strategy))
```


# Calculate the total number of steps taken per day
```{r , eval=TRUE, echo=TRUE}
total.steps.filled <- tapply(activity.data.filled$steps, activity.data.filled$date, FUN = sum, na.rm = TRUE)
summary(total.steps.filled)
```
# Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends
##  https://cran.r-project.org/web/packages/timeDate/index.html
```{r , eval=TRUE, echo=TRUE}
activity.data$day.type <- "weekday"
activity.data$day.type[isWeekend(activity.data$date)] <- "weekend"
activity.data$day.type <- as.factor(activity.data$day.type)
```

# Make a panel plot containing a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis). 
```{r, eval=TRUE, echo=TRUE}
pattern <- aggregate(steps ~ interval + day.type, data = activity.data, mean)
library(lattice)
xyplot(steps ~ interval | day.type, pattern, type = "l", layout = c(1, 2), xlab = "Interval", ylab = "Number of steps")
```

##  Calculate average steps of each 5-minutes interval for weekdays and weekend
```{r , eval=TRUE, echo=TRUE}
weekend <- pattern[which(pattern$day.type == "weekend"),] %>% group_by(day.type) %>% summarise(Steps = mean(steps))
weekend
weekday <- pattern[which(pattern$day.type == "weekday"),] %>% group_by(day.type) %>% summarise(Steps = mean(steps))
weekday
```


